---
title: "Can I see a doctor please?"
description: "Analysing the gender ratio of licensed doctors across UK countries"
date: 2018-03-24T23:24:14Z
draft: false
---

```{r, include = FALSE}
# TF 09/01/17
# Mapping the World with ggplot and open source natural earth data

# 
####
# Psuedocode ----
# 1. Load libraries
# 2. Load data and map
# 3. Clean data and map
# 4. Join data to map
# 5. Create unautorised absence map
# 6. Create authorised absence map
# Annex
####

####
# 1. Load libraries ----
library(leaflet)
library(geojsonio)
library(rgdal)
library(sp)
library(dplyr)
library(plyr)
library(data.table)
library(RColorBrewer)
library(raster)
library(ggplot2)
library(rgeos)
library(readr)
library(mapproj)
library(tictoc)
library(ggmap)
library(maps)
library(ggthemes)
library(htmlwidgets)
library(tidyr)
####
# 2. Load data and create shapefiles for maps ----

countries_shapefile <- shapefile("Doctor Data/ne_10m_admin_0_map_units/ne_10m_admin_0_map_units.shp")
# plot(countries_shapefile)


# One way to filter countries by starting letter of country!

countries_uk = subset(countries_shapefile, SUBUNIT %like% "England" | 
                        SUBUNIT%like% "Wales" |
                        SUBUNIT %like% "Scotland" |
                        SUBUNIT %like% "Northern Ireland")

countries_uk <- spTransform(countries_uk, CRS("+proj=longlat +ellps=WGS84"))

doctor_data <- read.csv("Doctor Data/doctors.csv")

doctor_data %>% 
  dplyr::filter(Year == 2017) %>% 
  dplyr::select(Country, Gender, Year, Number) %>%
  tidyr::spread(Gender, Number) -> doctor_data

doctor_data$Male <- as.numeric(sub(",", "", doctor_data$Male, fixed = TRUE))
doctor_data$Female <- as.numeric(sub(",", "", doctor_data$Female, fixed = TRUE))

doctor_data %>%
  mutate(majority_gender = ifelse(Male > Female, "Majority of doctors are male", 
                           ifelse(Female > Male, "Majority of doctors are female", "Gender Equality")))  %>%
  mutate(total = Male + Female) %>%
  mutate(Male_prop = (Male / total)*100) %>%
  mutate(Female_prop = (Female / total)*100) -> doctor_data

doctor_data$Male_prop <- format(round(doctor_data$Male_prop, 2), nsmall = 2)
doctor_data$Female_prop <- format(round(doctor_data$Female_prop, 2), nsmall = 2)

doctor_data$Country <- as.character(doctor_data$Country)

doctor_data <- droplevels(doctor_data)

doctor_data$majority_gender <- as.factor(doctor_data$majority_gender)

data_for_mapping <- sp::merge(countries_uk,
                          doctor_data,
                          by.x = 'SUBUNIT',
                          by.y = 'Country',
                          duplicateGeoms = TRUE)



map_pal = colorFactor(c('purple', '#4169e1'), data_for_mapping$majority_gender)





hoverText <- sprintf("<div style='font-size:12px;width:200px;float:left'>
            <span style='font-size:18px;font-weight:bold'>%s</span><br/> 
            <div style='width:95%%'>
              <span style='float:left'>Male</span>
                     <span style='float:right'>Female</span>
                     <br/>
                     <span style='color:black;float:left'>%s%%</span>
                     <span style='color:black;float:right'>%s%%</span><br clear='all'/>
                     <span style='background:#D4DCF7;width:%s%%;float:left'>&nbsp;</span>
                     <span style='background:#E7CCFC;width:%s%%;float:right'>&nbsp;</span>
                     </div>
                     <br/><span style='font-size:10px'>%s</span>
                     </div>",
                      data_for_mapping$SUBUNIT, 
                      data_for_mapping$Male_prop, data_for_mapping$Female_prop,
                      data_for_mapping$Male_prop, data_for_mapping$Female_prop,
                     data_for_mapping$majority_gender) %>%
  lapply(htmltools::HTML)




map <- leaflet(data_for_mapping,
  options=leafletOptions(attributionControl = FALSE, 
  dragging = FALSE, zoomControl = FALSE, minZoom = 5.8, maxZoom = 5.8)) %>%
  addPolygons(fillColor=~map_pal(data_for_mapping$majority_gender),
              weight = 1,
              label = ~hoverText,
              color = "grey",
              labelOptions = labelOptions(
                offset = c(-100,-140),
                #direction='bottom',
                textOnly = T,
                style=list(
                  'background'='rgba(255,255,255,0.95)',
                  'border-color' = 'rgba(0,0,0,1)',
                  'border-radius' = '4px',
                  'border-style' = 'solid',
                  'border-width' = '4px')),
              highlightOptions = highlightOptions(weight = 3, bringToFront = TRUE)) %>%
  setMaxBounds(lat1 = 59.2, lng1 = 10, lat2 = 49.2, lng2 = -10) %>%
  htmlwidgets::onRender(
    "function(el, t) {
    var myMap = this;
    // get rid of the ugly grey background
    myMap._container.style['background'] = '#ffffff';
    }") 
    


```

<div style= "float:right;position: relative; top: -10px;"> 
```{r, echo = FALSE, warning = FALSE, fig.height = 7.5, fig.width=5, fig.align="right"}
map
```
</div>


> Scotland is the first UK country to have more licensed female doctors than males. 

The demographics of the medical profession are changing. In 2017, latest data from the General Medical Council showed that the number of licensed doctors in Scotland was greater than those of men for the first time. 

Their annual report, [the state of medical education and practice 2017](https://www.gmc-uk.org/publications/somep2017.asp), showed that 51.01% of all licensed doctors in Scotland were female, relative to just 44.13% in Wales. 

This demographic shift highlights the changing medical profession in the UK. Currently XX% of doctors in training, the future of the UK's medical workforce, are female. This suggests that in the future, we are likely to see a female dominated medical profession in the UK.

However, some areas of specialism within medicine may take much longer to see such change. Surgery 

\n


\n

\n

\n

\n
